Verifying Demonstrations

### 2020-03-13 paper/formal-verification.tex

\section{Bit-by-bit distance-bounding Schnorr protocol}%

As mentioned in \cref{related-work}, while there exists a \iac{PPK} 
protocol~\cite{ProProx}, there is none for discrete logarithms. We will now introduce such a protocol to enable us to have \ac{DB} anonymous credentials.

\subsection{Reattempting a distance-bounding Schnorr protocol}%
\label{DBSchnorr-n-bit}

In the original \ac{DB} paper by \citet{DistanceBounding}, one of the distance-bounding protocols is based on the Schnorr identification scheme~\cite{Schnorr}.
However, this \ac{DB} Schnorr protocol was shown to be prone to distance hijacking~\cite{DistanceHijacking, TamarinDB}, in addition to not being secure against terrorist fraud.
We now propose another way to turn the Schnorr protocol into a public-key \ac{DB} protocol that is secure against \ac{DBMF}, \ac{DBDF}, \ac{DBDH} and \ac{DBTF}.
It is important in our setting that the protocol is a public-key protocol, both to reach strong privacy properties and also because this property implies protection against an impersonating verifier.

We present the protocol in \cref{SchnorrFigure}.
The (cyclic) group with generator \(g\) and order \(q\) are system parameters.
The private key \(\alpha\) with public key \(A = g^\alpha\) are generated once by the prover in the setup phase.

\begin{figure*}
  \centering
  \small
  \setlength{\ProtoArrowLength}{0.07\linewidth}
  \begin{tabular}{p{0.40\linewidth}cp{0.40\linewidth}}
    \(\DBSprove[g, q, \alpha, A = g^\alpha]\):
    & &
    \(\DBSverify[g, q, A]\):
    \\
    \midrule

    \multicolumn{3}{c}{\textbf{Setup}} \\

    \(\rho\rgets \ZZ_q, R\gets g^{\rho}\)
    & \(\ProtoSendRight{R}\)
    &
    \\

    % null
    & \(\ProtoSendLeft{c_0, c_1}\)
    & \(c_0\rgets \bin^k, c_1\rgets \bin^k\)
    \\

    \(s_0\gets \rho - c_0\alpha \pmod q\)
    &
    & \(b\rgets \bin\)
    \\

    \(s_1\gets \rho - c_1\alpha \pmod q\)
    &
    & Prepare \(B\in \bin^{|q|+l}\), with \(|q|\) bits set to \(b\) and the 
    remaining \(l\) bits to \(1-b\).
    \\

    \midrule
    \multicolumn{3}{c}{\textbf{Distance-bounding}: \(\forall i: 0\leq i < |q| + 
        l, j\gets 0\)} \\

%    % null
%    &
%    & Random delay
%    \\

    % null
    & \(\ProtoSendLeft{b_i}\)
    & \(b_i\gets B[i]\)
    \\

    \(r_i\gets s_{b_i}[jb_i + (i-j)(1-b_i)]\)
    & \(\ProtoSendRight{r_i}\)
    & Record \(\Delta t_i\)
    \\

    \(j \gets j + b_i\)
    &
    &
    \\

    \midrule
    \multicolumn{3}{c}{\textbf{Verification}}
    \\

    % null
    &
    & Construct \(s_b\) as the concatenation of \(r_i\)'s for which \(b_i = 
      b\).
      Accept if \(R = g^{s_b} A^{c_b}\).
    \\

  \end{tabular}
  \caption{%
    One-round protocol instance of the \(\DBSprove\leftrightarrow \DBSverify\) \ac{DB} Schnorr protocol for \(\PK[\alpha][A = g^\alpha]\).
    The protocol should be repeated in full to achieve the desired soundness 
    and distance-bounding errors.
  }%
  \label{SchnorrFigure}
\end{figure*}

During one round, the prover commits to a random nonce: more precisely he chooses \(\rho\rgets \ZZ_q\) uniformly at random, computes \(R\gets g^\rho\) and sends \(R\) to the verifier.
The verifier generates two challenges \(c_0\rgets \bin^k, c_1\rgets \bin^k\) and sends them to the prover.
The prover computes \(s_0\gets \rho - c_0\alpha, s_1\gets \rho - c_1\alpha\).
This step is the main difference to the original Schnorr protocol: the verifier selects two challenges and the prover computes two responses --- but still only \emph{one nonce}, \(\rho\).
This is also different from the original Brands-Chaum protocol, in which the 
prover and verifier jointly construct \emph{one} challenge with \emph{one} 
response.

We let \(|q| = \ceil{\log_2 q}\) denote the length of \(q\) in bits.
The verifier will request all \(|q|\) response-bits from one challenge (say 
\(s_b\)) and only \(0 < l\leq |q|\) from the other (\(s_{1-b}\)).
Then the verifier can authenticate the prover by checking if \(R = 
  g^{s_b}A^{c_b}\).

\subsection{Security analysis}

The prover must know the responses for both challenges to successfully pass the \ac{DB} phase.
The reason for the two challenges but only one random nonce is that knowing the responses to both means learning \(\alpha\).
This allows us to prevent \ac{DBTF}.
Bundling the authentication into the distance-bounding phase (difference from Brands-Chaum) prevents \ac{DBDH}, and both adaptations together aim at \ac{DBMF} resistance. %reducing success of MF according to number of repetitions.
The reason we use secret sharing and not \(s_b\) directly, is to bound the \(l\) bits the verifier must learn of \(s_{1-b}\).
Without secret sharing, the verifier learns \(l\) bits of \(s_{1-b}\), but with secret sharing the verifier learns \(l\) bits of \emph{one share}.
Thus, we still have information theoretic security for \(s_{1-b}\)
although we reveal the \(l\) bits. (See \cref{l-bit-problem} for an
  in-depth treatment of what we call the \(l\)-bit problem.)

The intuition behind the protocol security is as follows.
To achieve malicious-verifier zero-knowledge, choose \(k\) logarithmic in the security parameter \(\lambda\) and repeat the protocol \(n\) times, such that 
the knowledge error, \(2^{-kn}\), becomes small enough.
Repeating the protocol is also needed to decrease the success probability of some of adversaries presented below.

The \ac{DB} phase protects against distance fraud.
Once the prover has received \(l+1\) challenges set to \(b\), it knows that the remaining challenges must also be \(b\).
\Iac{DBDF} prover must thus wait for the challenge bit \(b_i\) before responding with \(r_i\) for \(l+1\) challenges.
Thus, the probability of successfully guessing the order of the challenge bits is \(2^{-(l+1)}\) per round or, in total, \(2^{-(l+1)n}\).

The \ac{DB} phase also ensures that \iac{DBMF} adversary will fail the \ac{DB} phase for at least one round.
The verifier will send \(|q| + l < 2|q|\) challenges.
By replying, the prover reveals \(|q|+l\) bits distributed over \(s_0, 
  s_1\).
\emph{We assume} that learning \(\ceil{(|q|+l)/2}\) bits of \(s_0\) and 
\(\floor{(|q|+l)/2}\) bits of \(s_1\) yields no advantage for the adversary.
Now, the adversary has two options.
First, the adversary can buffer \(l\) bits of both \(s_0^{(0)}\concat s_0^{(1)}, s_1^{(0)}\concat s_1^{(1)}\) by requesting them from the prover.
However, at this point the adversary must wait for the challenge bit \(b_{l+1}\) from the verifier and then relay that challenge to the prover to receive the correct \(r_i\) --- this relay will be detected.
The alternative is to guess \(b\) and extract \(s_b\) for this challenge from the prover.
Guessing \(b\) yields \(1/2\) probability that the \ac{DBMF} adversary will guess it correctly.
Since we must execute the protocol \(n\) times, the probability to pass all these rounds is reduced to \(2^{-n}\).
If the adversary requests \(s_b\) in full, the adversary can also guess the 
remaining \(|q|-l\) bits of \(s_{1-b}\).
Based on our assumption above, the probability of guessing these bits 
successfully is \(2^{-(|q|-l)}\).
Thus the success probability of the \ac{DBMF} adversary is \emph{at most} 
\(2^{-n}+2^{-(|q|-l)n}\).

The protocol is \ac{DBTF}-resistant. 
Indeed, if the malicious prover gives both responses to the accomplice, the accomplice can compute his secret key.
The probability of success of \ac{DBTF} is thus reduced to guessing \(b\), \ie \(1/2\) per round or, in total, \(2^{-n}\).

The protocol is also secure against distance hijacking due to the fact that the authenticated bitstring is used during the \ac{DB} phase, not the challenge bitstring as it is the case for the original protocol of Brands-Chaum.
Thus the probability of success is equivalent of a collision for the responses 
for the chosen challenge --- in each round.

Finally, it follows from the original Schnorr protocol that this is still \iac{PK} and \ac{ZK}.
\daniel{I should prove this for the composed protocol, I should check
  the details with Douglas.}

Besides the security against the standard attacks on distance-bounding
protocols, we also need to ensure that our modification does not leak
any information to the adversary via the aforementioned \(l\) bits.

\subsection{The \(l\)-bit problem}
\label{l-bit-problem}

We will now systematically investigate the concern of the adversary
gaining any advantage from learning \(l\) bits from one of the two challenges.

\paragraph*{A special case}

We will start with a special case of the \(l\)-bit problem.
For a session~\(i\), we have two possible responses to the two challenges:
\(s_i = r_i + c\alpha\) and \(s_i' = r_i + c'\alpha\).
Let \(c = 1\) and \(c' = 0\).
We assume that the adversary learns all \(m\) bits of \(s_i\), but only \(l\) 
bits of \(s_i'\).
This special case yields the linear equation system
\begin{align*}
  s_i &= r_i + \alpha
  = s_i'[m-1] 2^{m-1} + \dotsb + s_i'[1] 2^1 + s_i'[0] 2^0 + \alpha &\pmod q\\
  s_i'[j_1] &= b_{j_1} &\pmod 2\\
  s_i'[j_2] &= b_{j_2} &\pmod 2\\
            &\vdots \\
  s_i'[j_l] &= b_{j_l}, &\pmod 2
\end{align*}
where \(s_i'[j]\) denotes the \(j\)th bit of \(s_i'\), and \(j_1, \dotsc, 
j_l\in \{0, \dotsc, m-1\}\) are indices for the bits chosen by the adversary 
and \(b_{j_1}, \dotsc, b_{j_l}\in \bin\) are the bit values.

\begin{lemma}
  Assume the special case given above.
  If \(l < m+\frac{1}{n}-1\), then the equation system is underdetermined and 
  we cannot solve for \(\alpha\).
\end{lemma}
\begin{proof}
  The equation system above yields \(l+1\) equations and \(m+1\) variables.
  If we repeat the protocol \(n\) times, we get \(n\) such sessions.
  This yields an equation system of \(n(l+1)\) equations and \(nm+1\) variables 
  (\(\alpha\) is not repeated).
  If \(n(l+1) < nm+1\), this system is underdetermined and cannot be solved.
  After some rewriting we find that
  \begin{align*}
    n(l+1) &< nm+1 \\
    l+1 &< m + \frac{1}{n} \\
    l &< m + \frac{1}{n} -1.
  \end{align*}
  Which concludes the proof.
\end{proof}

The number of possible solutions is vital for us, since that is an estimate of 
how difficult it is for the adversary to guess \(\alpha\).

\begin{lemma}\label{simple-l-bit-advantage}
  Assume the special case above.
  There are \(m-l\) free variables, each taking a value from \(\bin\).
  This yields \(2^{m-l}-1\) unique solutions for each session.
\end{lemma}
\begin{proof}
  The result directly follows from the fact that there are \(m-l\) free 
  variables.
  Each free variable can only take two possible values (0 or 1), consequently 
  there are \(2^{m-l}-1\) unique solutions.
\end{proof}

\paragraph*{Arbitrarily chosen \(c\neq c'\)}

For any choice of \(c\) and \(c'\) such that \(c\neq c'\), we have
\begin{align}
  \label{eqn:general-c}
  s_i &= r_i + c\alpha &\pmod q\\
  \label{eqn:general-cprime}
  s_i' &= r_i + c'\alpha. &\pmod q
\end{align}
We can transform this system of equations into our special case above by 
Gaussian elimination: add \cref{eqn:general-c} multiplied by \(\frac{c'}{c}\) 
to \cref{eqn:general-cprime}.
The multiplicative inverse \(\frac{1}{c}\) exists since \(q\) is prime.
This proves the following lemma.

\begin{lemma}
  Any choice of \(c, c'\) can be turned into our special case of \(c = 1, c' = 
  0\).
\end{lemma}

This means that \cref{simple-l-bit-advantage} is also true for the general case.

\paragraph*{The unconditional \(l\)-bit problem}

We run the protocol \(n\) times and let the adversary learn \(m+l\) bits of the 
result~\(s_i, s_i'\).
This yields the following equation system:
\begin{align*}
  s_0[0]2^0 + \dotsb + s_0[m-1]2^{m-1} &= r_0 + c\alpha &\pmod q\\
  s_0'[0]2^0 + \dotsb + s_0'[m-1]2^{m-1} &= r_0 + c'\alpha &\pmod q\\
  \vdots \\
  s_{n-1}[0]2^0 + \dotsb + s_{n-1}[m-1]2^{m-1} &= r_{n-1} + c\alpha &\pmod q\\
  s_{n-1}'[0]2^0 + \dotsb + s_{n-1}'[m-1]2^{m-1} &= r_{n-1} + c'\alpha, &\pmod 
  q
\end{align*}
where \(m+l\) of \(s_i[j]\)'s and \(s_i'[j']\)'s for each \(i\) are known by 
the adversary.

\begin{lemma}\label{l-bit-advantage}
  If \[l < m-1+\frac{1}{n}-\frac{f(\lambda)}{n},\] then there are at least 
  \(2^{f(\lambda)}-1\) unique solutions, for some function~\(f\) of the 
  security parameter~\(\lambda\).
\end{lemma}
\begin{proof}
  Consider one instance of the protocol:
  \begin{align*}
    s_i[0]2^0 + \dotsb + s_i[m-1]2^{m-1} &= r_i + c\alpha \\
    s_i'[0]2^0 + \dotsb + s_i'[m-1]2^{m-1} &= r_i + c'\alpha.
  \end{align*}
  This yields \(2m+2\) variables with two equations, thus \(2m\) free variables.
  Let the adversary learn \(m+l\) of the bits, \ie left-hand side \(s_i[j]\)'s or 
  \(s_i'[j']\)'s.
  This leaves \(2m-(m+l) = m-l\) free variables and \(2^{m-l}-1\) possible 
  solutions.

  Repeat this \(n\) times.
  This yields \(n(2m+1)+1\) variables and \(2n\) equations, thus \(2nm+n+1-2n = 
  2mn - n +1 = n(2m-1)+1\) free variables.
  Let the adversary learn \(m+l\) bits in each round, thus \(n(m+l)\) in total.
  This yields \(n(2m-1)+1 -n(m+l) = 2nm-n+1-nm-nl = nm-n+1-nl = n(m-1-l)+1\) 
  free variables and \(2^{n(m-1-l)+1}-1\) unique solutions.
  Consequently, if we want at least \(2^{f(\lambda)}-1\) unique solutions, we 
  have that
  \begin{equation*}
    nm-n-nl+1 > f(\lambda) \iff m-1+\frac{1}{n}-\frac{f(\lambda)}{n} > l.
  \end{equation*}
  Which concludes the proof.
\end{proof}

\paragraph*{The \(l\)-bit problem with commitments}

The adversary has a commitment to the randomness~\(r_i\) in the form of \(R_i = 
g^{r_i}\) and a commitment to \(\alpha\) in the form of \(A = g^\alpha\).
This means that
\begin{align*}
  R = A^{-c'} g^{s'} &= A^{-c'} g^{s'[0] 2^0 + \dotsb + s'[m-1] 2^{m-1}} \\
                     &= A^{-c'} g^{s'[0] 2^0} \dotsb g^{s'[m-1] 2^{m-1}} \\
                     &= A^{-c'} g^{s'_{\text{known}}} g^{s'_{\text{unknown}}}.
\end{align*}
We will now show how the hardenss of finding \(s'_\text{unknown}\) relates to 
the hardness of finding discrete logarithms.

\begin{definition}[\Acl*{DLP}]
  Let \(g\) be a generator for a group of prime order~\(q\).
  Given \(g, q, A = g^\alpha\), compute \(\alpha\).
\end{definition}

\begin{definition}[DB-Schnorr Problem]%
  \label{DBSchnorr-problem}
  Let \(g\) be a generator for a group of prime order~\(q\).
  Give \(g, q, R = g^\rho A = g^\alpha\) to the adversary.
  Let the adversary choose \(c\) and additionally give him \(\hat s\) such that 
  \(\hat s\) are the \(l\) \emph{least significant} bits of \(s = \rho + 
  c\alpha \pmod q\).
  The adversary wins if he returns \(\hat s'\) such that \(\hat s + \hat s' = s 
  \pmod q\).
\end{definition}

\begin{lemma}
  Given an algorithm~\(\algA_{\DBS}\) that solves the DB-Schnorr Problem 
  (\cref{DBSchnorr-problem}) with probability \(\epsilon\), we can construct an 
  algorithm \(\algA_{\text{DL}}\) that solves the \ac{DLP} with probability 
  \(\epsilon 2^{-(l+k)}\).
\end{lemma}

\begin{proof}
  We will now construct the algorithm \(\algA_{\text{DL}}\) given the algorithm 
  \(\algA_{\DBS}\) and the \ac{DLP} challenge \(g, q, A_{\text{DL}}\).
  Assume that \(\algA_{\DBS}\) succeeds with probability \(\epsilon\).

  We proceed as follows.
  Guess \(c\rgets \ZZ_k\), choose \(\hat s\rgets \ZZ_{2^l}\).
  Let \(\hat R\gets A_{\DBS}^c A_{\text{DL}} g^{\hat s}\) and give
  \(\hat R\) to \(\algA_{\DBS}\).
  If \(c'\gets \algA_{\DBS}(g, q, A_{\DBS}, \hat R)\) returns \(c'\neq c\), 
  abort.
  The probability of success is \(2^{-k}\).
  Otherwise, give \(\hat s\) to \(\algA_{\DBS}\).
  Now, by construction, we have that \(
    \hat R = A_{\DBS}^c g^{\hat s} A_{\text{DL}}.
  \)

  If \(A_{\text{DL}} = g^{\alpha_{\text{DL}}}\) is chosen such that 
  \(\alpha_{\text{DL}}\) has the \(l\) least significant bits set to zero 
  (probability \(2^{-l}\)), then will have \(
    \alpha_{\text{DL}}\gets \algA_{\DBS}(g, q, A_{\DBS}, \hat R, \hat s)
  \) with probability \(\epsilon\).
  Thus, the total probability of success is \(
    \epsilon 2^{-l} 2^{-k} = \epsilon 2^{-(l+k)}.
  \)
\end{proof}

As mentioned above, both \(l\) and \(k\) are logarithmic in the security 
parameter.

\endinput
\subsection{Formal Verification}

>>>>

\section{Formal Verification}
